---
output: github_document
---

# Drost 1967 "Töpferei in Afrika"

- - -

> :warning: **Warning:** This repository is still work-in-progress!!

- - -

This repository contains geo-referenced information on potters' communities throughout Africa.

The maps were geo-references using QGIS (v3.22.1-Białowieża) relying on nine manually set reference points (tranformations parameters: Polynomial 3 & cubic spline) into EPSG:4326 CRS.

## Stages of the *chaîne opératoire* 

The following stages of the *chaîne opératoire* were recorded:

| Stage       | Description   | Source |
| ----------- | ------------- | ------ |
| RAWMATERIAL | Raw material sources (fluvial/river, termite mounds, mining) | p. 19 map 1 |
| PREPARATION | Raw material preperation | p. 33 map 2 |
| PRIMARY_SHAPING | Primary shaping technique | p. 55-79 map 3-6 |
| SECONDARY_SHAPING | Secondary shaping technique | p. 55-79 map 3-6 |
| FIRING      | Firing        | p. 217 map 17 |

```{r setup, include=FALSE}
library(ggplot2)
library(rnaturalearth)
library(tidyverse)
library(geojsonsf)

d <- geojsonsf::geojson_sf("Drost1967.geojson")

# naturalearth geodata
land10 <- ne_download(scale = 10, type = 'land', category = 'physical', returnclass = "sf")
rivers10 <- ne_download(scale = 10, type = 'rivers_lake_centerlines', category = 'physical', returnclass = "sf")
lakes10 <- ne_download(scale = 10, type = 'lakes', category = 'physical', returnclass = "sf")
coast10 <- ne_download(scale = 10, type = 'coastline', category = 'physical', returnclass = "sf")
```

## Maps


```{r map, warning = FALSE, echo = FALSE, fig.align = 'center', dpi = 300, out.width = '100%'}
cowplot::plot_grid(
  ggplot() +  
    geom_sf(data = land10, fill = "white", color = NA) + 
    geom_sf(data = coast10) + 
    geom_sf(data = rivers10) +
    geom_sf(data = lakes10) + 
    geom_sf(data = d %>% dplyr::filter(!is.na(RAWMATERIAL)), aes(fill = RAWMATERIAL), shape = 21) + 
    coord_sf(xlim = c(-18, 50),
             ylim = c(-35, 35)) + 
    theme(legend.position = "right"),
  ggplot() +  
    geom_sf(data = land10, fill = "white", color = NA) + 
    geom_sf(data = coast10) + 
    geom_sf(data = rivers10) +
    geom_sf(data = lakes10) + 
    geom_sf(data = d %>% dplyr::filter(!is.na(PREPARATION)), aes(fill = PREPARATION), shape = 21) + 
    coord_sf(xlim = c(-18, 50),
             ylim = c(-35, 35)) + 
    theme(legend.position = "right"),
  ggplot() +  
    geom_sf(data = land10, fill = "white", color = NA) + 
    geom_sf(data = coast10) + 
    geom_sf(data = rivers10) +
    geom_sf(data = lakes10) + 
    geom_sf(data = d %>% dplyr::filter(!is.na(PRIMARY_SHAPING)), aes(fill = PRIMARY_SHAPING), shape = 21) + 
    coord_sf(xlim = c(-18, 50),
             ylim = c(-35, 35)) + 
    theme(legend.position = "right"),
  ggplot() +  
    geom_sf(data = land10, fill = "white", color = NA) + 
    geom_sf(data = coast10) + 
    geom_sf(data = rivers10) +
    geom_sf(data = lakes10) + 
    geom_sf(data = d %>% dplyr::filter(!is.na(FIRING)), aes(fill = FIRING), shape = 21) + 
    coord_sf(xlim = c(-18, 50),
             ylim = c(-35, 35)) + 
    theme(legend.position = "right"), 
  align = "hv"
)
```

## Collective *chaînes opératoires*

```{r echo = FALSE}
d.clean <- d %>% 
  as.data.frame() %>%
  dplyr::select(RAWMATERIAL, PREPARATION, PRIMARY_SHAPING, SECONDARY_SHAPING, FIRING) %>%
  stats::na.omit() %>%
  tibble::rowid_to_column("ID")

# run MCA for grouping

d.mca.res <- FactoMineR::MCA(d.clean %>% tibble::column_to_rownames("ID"), graph = F)

factoextra::fviz_mca_biplot(d.mca.res, repel = T)

d.mca.clust <- FactoMineR::HCPC(d.mca.res, nb.clust = 0, graph = FALSE, iter.max = 100, min = 3, max = NULL)

# d.mca.clust$desc.var

d.clust <- d.mca.clust$data.clust %>%
  tibble::rownames_to_column("ID") %>% 
  dplyr::mutate(ID = as.integer(ID)) %>% 
  dplyr::select(ID, clust) # %>% dplyr::left_join(d.clean, by = "ID")

# building edgelist per community


```




```{r eval=FALSE, echo=FALSE}
d.lng <- d.clean %>% 
  reshape2::melt(id.vars = "ID") %>%
  dplyr::arrange(ID)

c.unq <- unique(d.clean$ID)
el.lst <- list()
v.n.lst <- list()
for (i in 1:length(c.unq)) {
  i = 1
  
  # c.el <- 
  d.lng %>% 
    dplyr::filter(ID == c.unq[i])
  
  
    %>%
    dplyr::mutate(FROM = lag(OPERATION), 
                  TO = OPERATION) %>%
    dplyr::select(OBJ, FROM, TO) %>%
    dplyr::filter(!is.na(FROM))
  
  
  
  
  v.n <- network::network(
      v.el[,c("FROM", "TO")], 
      loops = TRUE, 
      multiple = TRUE)
  
  
  v.n.df <- ggnetwork::ggnetwork(v.n) %>%
    dplyr::mutate(OBJ = unique(v.el$OBJ))

  el.lst[[i]] <- v.el
  v.n.lst[[i]] <- v.n.df
}
el <- do.call(rbind, el.lst)
el

v.n <- do.call(rbind, v.n.lst)
v.n


```

